---
title: "Activity Assignment"
author: "Baird"
date: "January 28, 2017"
output:
  html_document: default
  pdf_document: default
---
## Assignment objectives
###Commit containing full submission

####-1)Code for reading in the dataset and/or processing the data

####-2)Histogram of the total number of steps taken each day

####-3)Mean and median number of steps taken each day

####-4)Time series plot of the average number of steps taken

####-5)The 5-minute interval that, on average, contains the maximum number of steps

####-6)Code to describe and show a strategy for imputing missing data

####-7)Histogram of the total number of steps taken each day after missing values are imputed

####-8)Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

####-9)All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

#####load data from activity.csv file

```{r}
#quartz()
#dev.off()
#plot.new()
library(ggplot2)
```
####-1)Code for reading in the dataset and/or processing the data
```{r}
activity <- read.csv("activity.csv")
activity$date2<-as.Date(as.character(activity$date), '%Y-%m-%d')
activity_ag <- aggregate(steps~date2, data=activity, FUN=sum,na.rm=TRUE)
activity_ag<-activity_ag[order(activity_ag$date2),]
```
####-2)Histogram of the total number of steps taken each day
```{r}
#plot.new()
#g<-g+ggtitle("This is a longer\ntitle than expected")
ggplot(activity_ag) + geom_col(aes(x=date2,y=steps,group=date2,fill=steps)) +
  xlab("Dates Range for Steps Collection")+
  ylab("Total Steps per Day")+
  ggtitle("Plot of Total Steps Taken Per Day \nDuration-Recorded Date Interval")+
  theme(plot.margin=unit(c(0,0,0,0),"mm"))
#dev.off()
```
####-3)Mean and median number of steps taken each day
```{r}
activity_int <- aggregate(steps ~ interval, data=activity, FUN=mean)

print(paste0("Mean steps taken over interval: ", steps_mean<-mean(activity_ag$steps)))

print(paste0("Median steps taken over interval: ", steps_median<-median(activity_ag$steps)))
```
####-4)Time series plot of the average number of steps taken
```{r}
activity_interval_ag <- aggregate(steps ~ interval, activity, mean, rm.na=T)

plot(x = activity_interval_ag$interval,y = activity_interval_ag$steps,
        type = "l",lwd=3, col="dark green",
        main = "5 Minute Intervals \n Average Number of Steps Taken Across Study Interval",
        xlab = "5 Minute Intervals Averages (Mean)",
        ylab = "Average Number of Steps Taken,\n Averaged Across Study Interval")
#dev.off()
```
####-5)The 5-minute interval that, on average, contains the maximum number of steps
```{r}
print(paste0("The 5-minute interval that contains the maximum number of steps: ",
activity_interval_ag$interval[which.max(activity_interval_ag$steps)]))

sum(is.na(activity))

#create new data group for adding fake data to missing steps
activity_merge<-activity

#create a function to return steps mean from time interval
step_inteveral <- function(int){
  activity_interval_ag$steps[activity_interval_ag$interval==int]
}
#testing function over missing data
#step_inteveral(activity_merge$interval[is.na(activity_merge$steps)])

#replace na data with mean steps
activity_merge$steps[is.na(activity_merge$steps)]<- round(as.numeric(lapply(activity_merge$interval[is.na(activity_merge$steps)],step_inteveral), digits = 0))
#activity_merge$steps[is.na(activity_merge$steps)]<- round(lapply(activity_merge$interval[is.na(activity_merge$steps)],step_inteveral), digits = 0)

activity_merge_ag <- aggregate(steps~date2, data=activity_merge, FUN=sum,na.rm=TRUE)
activity_merge_ag<-activity_merge_ag[order(activity_merge_ag$date2),]
```
####-6)Code to describe and show a strategy for imputing missing data
```{r}
#plot.new()
ggplot(activity_merge_ag) + geom_col(aes(x=date2,y=steps,group=date2,fill=steps)) +
  xlab("Dates Range for Steps Collection (Imputed Data)")+
  ylab("Total Steps per Day")+
  ggtitle("Plot of Total Steps Taken Per Day \nDuration-Recorded Date Interval")+
  theme(plot.margin=unit(c(0,0,0,0),"mm"))

activity_merge_int <- aggregate(steps ~ interval, data=activity_merge, FUN=mean)

print(paste0("Mean steps taken over supplemented (imputed) interval: ", steps_merge_mean<-mean(activity_merge_ag$steps)))

print(paste0("Median steps taken over supplemented (imputed) interval: ", steps_merge_median<-median(activity_merge_ag$steps)))
```
####-7)Histogram of the total number of steps taken each day after missing values are imputed
```{r}
activity_interval_merge_ag <- aggregate(steps ~ interval, activity_merge, mean, rm.na=T)

activity_weekends<-activity[weekdays(as.Date(activity$date)) %in% c("Saturday", "Sunday"),]
activity_weekdays<-activity[weekdays(as.Date(activity$date)) %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"),]

activity_int_weekend_ag <- aggregate(steps ~ interval, activity_weekends, mean, rm.na=T)
activity_int_weekday_ag <- aggregate(steps ~ interval, activity_weekdays, mean, rm.na=T)
par(mfrow = c(2, 1))
plot(x = activity_int_weekend_ag$interval,y = activity_int_weekend_ag$steps,
        type = "l",lwd=3, col="dark green",
        main = "5 Minute Intervals \n Average Number of Steps Taken Across Study Interval\nWeekends",
        xlab = "5 Minute Intervals Averages (Mean)",
        ylab = "Steps")

plot(x = activity_int_weekday_ag$interval,y = activity_int_weekday_ag$steps,
        type = "l",lwd=3, col="blue",
        main = "5 Minute Intervals \n Average Number of Steps Taken Across Study Interval\nWeekdays",
        xlab = "5 Minute Intervals Averages (Mean)",
        ylab = "Steps")
